### 2022-07-14 chapters/optimization/python-dir/python.tex
update:  introduction of python

%%%%%%%%%%% READ AGAIN 
articular, Noureddine in 2012 and Pereira in 2017 conducted empirical power measurements on this topic, and both concluded that compiled programming languages consume less power than dynamic ones.They found that a programming language that is interpreted, like Python, can use up to 7,584 times more energy than C. As a result, in this chapter, we will look at the obvious optimizations that may be done to Python legacy programs to lower their energy footprint.Since Python is used by many software services in both public and private cloud infrastructures, we think that our contributions will help a wide range of legacy systems by reducing ICT carbon emissions and lowering cloud bills for the resources used by these services. This chapter focuses on runtime improvements that developers can employ to reduce the power consumption of Python applications.We start by looking at how the choices made by programmers, like the type of data structure or loop, affect how much energy the code as a whole uses. Then we discuss additional topics, such as the level of concurrency, and so on.We will discuss another non-intrusive technique to optimize this code's energy consumption.Alternative interpreters and libraries dedicated to optimizing code without affecting its structures, such as ahead-of-time (AOT) compilation and just-in-time (JIT) libraries maintained by the community, are examples of these optimizations. 
% However, the operational conditions under which these measurements may threaten the validity of their results.

@@

Nowdays, Python seems to attract a large community of developers interested in data analysis, web development, system administration, machine learning---according to a survey conducted in 2018 by JetBrains---\footnote{\url{https://www.jetbrains.com/research/python-developers-survey-2018/}}one can fear that the wide adoption of dynamic programming languages, like Python, in production may critically hamper the power consumption of ICT.

As the popularity of such dynamic programming languages partly builds on the wealth and the diversity of their ecosystem (\emph{e.g.}, the NumPY, SciKit\,Learn, and Panda libraries in Python), one cannot reasonably expect that developers will likely move to an alternative programming language mostly for energy considerations.
Rather, we believe that a better option consists in leveraging the strength of this rich ecosystem to promote energy-efficient solutions in order to improve the power consumption of legacy software systems.

@@

the simplicity, flexibility and the clarity of Python code led this programming language to a huge success. Therefore, a larger community and a wide range of support and libraries that allowed it to touch several fields like data analyses, machine learning, natural language processing, web development, astrophysics etc.
\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/hannoiimplementation.png}
    \caption{Energy consumption of a recursive implementation of Tower of Hanoi program in different languages~\cite{noureddine_preliminary_2012}}
    \label{fig:hannoi}
\end{figure}

However, and according to \cite{pinto_energy_2017} and \cite{noureddine_preliminary_2012}. Python tends to be one more energy hungry . In figure \ref{fig:hannoi}, python consumes 30 times more than C or C++. The test was done on an implementation of \fnurl{hannoi test}{ https://en.wikipedia.org/wiki/Tower_of_Hanoi}  30 disks.

Python consumes a lot of energy mainly because it is slow in execution. Its Flexibility and simplicity caused it to drop off in performances, because Python gains its flexibility from being a dynamic language, therefore it needs an interpreter to execute its programs, which makes them much slower compared to the other ones that are written in compiled programming languages such as C and C++ or semi-compiled languages like Java.

As shown in Table \ref{fig:clbg}, one can see that in the most of the test cases Python takes more time to execute -the only case that he wasn't the last one was in the benchmark regx-redux where he beaten Go-, and in some cases the gab was huge such as in the n-body test where python took around 100 times more than C++.

\begin{table}[hbt]
    \begin{tabular}{l|*{5}c}
                           & C    & C++   & Java & python & Go    \\
        \hline
        pidigits           & 1.75 & 1.89  & 3.13 & 3.51   & 2.04  \\
        reverse-complement & 1.75 & 2.95  & 3.31 & 16.76  & 4.00  \\
        regx-redux         & 1.45 & 1.66  & 10.5 & 15.56  & 28.69 \\
        k-nucleotide       & 5.07 & 3.66  & 8.66 & 79.79  & 15.36 \\
        binary-trees       & 2.55 & 2.63  & 8.28 & 92.72  & 28.90 \\
        fasta              & 1.32 & 1.33  & 2.32 & 62.88  & 2.07  \\
        Fannkuch-redux     & 8.72 & 10.62 & 17.9 & 547.23 & 17.82 \\
        n-body             & 9.17 & 8.24  & 22.0 & 882.00 & 21.00 \\
        spectral-norm      & 1.99 & 1.98  & 4.27 & 193.86 & 3.95  \\
        Mandelbort         & 1.64 & 1.51  & 6.96 & 279.68 & 5.47
    \end{tabular}
    \caption{Comparison of the execution time between different programming languages using the computer language benchmarks game \cite{noauthor_which_nodate}}
    %  TODO : ADD reference 
    \label{fig:clbg}
\end{table}

\subsection{use cases}
to reduce the energy consumption of python we started by targeting the main usage of this programming language, wich is revealed to be data sciences and web developpement. Figure~\ref{fig:usecase} illustrate a study made by the jetbrain company on python developers \footnote{\fnurl{https://www.jetbrains.com/lp/python-developers-survey-2020}}. In this survey, multiple multiple answers were accepted.
\begin{figure}[hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_use_cases}
    \caption{use cases of python }
    \label{fig:usecase}
\end{figure}

\subsection*{Goal}
To reduce the energy consumption of python programs, we decided to start with studying its behaviour during the most common use cases. After this we will targtet the structures of the language itself in the hope to find some generic guidelines for more generic purpose as \citeauthor{hasan_energy_2016-1} did in their paper \cite{hasan_energy_2016-1}. After this we will measure the energy consumption of different python implementation for the goal of finding a non intrusive way to enhance this energy efficiency.

>>>>

%%%%%%%%%%% READ AGAIN 
% articular, Noureddine in 2012 and Pereira in 2017 conducted empirical power measurements on this topic, and both concluded that compiled programming languages consume less power than dynamic ones.They found that a programming language that is interpreted, like Python, can use up to 7,584 times more energy than C. As a result, in this chapter, we will look at the obvious optimizations that may be done to Python legacy programs to lower their energy footprint.Since Python is used by many software services in both public and private cloud infrastructures, we think that our contributions will help a wide range of legacy systems by reducing ICT carbon emissions and lowering cloud bills for the resources used by these services. This chapter focuses on runtime improvements that developers can employ to reduce the power consumption of Python applications.We start by looking at how the choices made by programmers, like the type of data structure or loop, affect how much energy the code as a whole uses. Then we discuss additional topics, such as the level of concurrency, and so on.We will discuss another non-intrusive technique to optimize this code's energy consumption.Alternative interpreters and libraries dedicated to optimizing code without affecting its structures, such as ahead-of-time (AOT) compilation and just-in-time (JIT) libraries maintained by the community, are examples of these optimizations. 
% However, the operational conditions under which these measurements may threaten the validity of their results.

@@

Nowadays, Python seems to attract a large community of developers who are interested in data analysis, web development, system administration, and machine learning—according to a survey conducted in 2018 by JetBrains ---\footnote{\url{https://www.jetbrains.com/research/python-developers-survey-2018/}}one can fear that the wide adoption of dynamic programming languages, like Python, in production may critically hamper the power consumption of ICT.
As the popularity of such dynamic programming languages partly builds on the wealth and the diversity of their ecosystem (\emph{e.g.}, the NumPY, SciKit\,Learn, and Panda libraries in Python), one cannot reasonably expect that developers will likely move to an alternative programming language mostly for energy considerations.
Rather, we believe that a better option consists of leveraging the strength of this rich ecosystem to promote energy-efficient solutions in order to improve the power consumption of legacy software systems.

@@

In the other hand,because of the interpreted nature of Python code, this programming language has paid a high price in terms of performance.  \ref{fig:clbg}, memory and energy consumption.

\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/hannoiimplementation.png}
    \caption{Energy consumption of a recursive implementation of Tower of Hanoi program in different languages~\cite{noureddine_preliminary_2012}}
    \label{fig:hannoi}
\end{figure}

According to \cite{pinto_energy_2017} and \cite{noureddine_preliminary_2012}, Python tends to be one more energy hungry . As onne can notice in figure \ref{fig:hannoi}, python consumes 30 times more than C or C++. The test was done on an implementation of \fnurl{hannoi test}{ https://en.wikipedia.org/wiki/Tower_of_Hanoi}  30 disks.

Python consumes a lot of energy, mainly because it is slow in execution. Its flexibility and simplicity caused it to drop off in performance, because Python gains its flexibility from being a dynamic language. Therefore, it needs an interpreter to execute its programs, which makes them much slower compared to the others that are written in compiled programming languages such as C and C++ or semi-compiled languages like Java.

As shown in Table\ref{fig:clbg}, one can see that in most of the test cases, Python takes more time to execute-the only case that he wasn't the last one was in the benchmark regx-redux where he beat Go-, and in some cases the gap was huge, such as in the n-body test where Python took around 100 times more than C++.
\begin{table}[hbt]
    \begin{tabular}{l|*{5}c}
                           & C    & C++   & Java & python & Go    \\
        \hline
        pidigits           & 1.75 & 1.89  & 3.13 & 3.51   & 2.04  \\
        reverse-complement & 1.75 & 2.95  & 3.31 & 16.76  & 4.00  \\
        regx-redux         & 1.45 & 1.66  & 10.5 & 15.56  & 28.69 \\
        k-nucleotide       & 5.07 & 3.66  & 8.66 & 79.79  & 15.36 \\
        binary-trees       & 2.55 & 2.63  & 8.28 & 92.72  & 28.90 \\
        fasta              & 1.32 & 1.33  & 2.32 & 62.88  & 2.07  \\
        Fannkuch-redux     & 8.72 & 10.62 & 17.9 & 547.23 & 17.82 \\
        n-body             & 9.17 & 8.24  & 22.0 & 882.00 & 21.00 \\
        spectral-norm      & 1.99 & 1.98  & 4.27 & 193.86 & 3.95  \\
        Mandelbort         & 1.64 & 1.51  & 6.96 & 279.68 & 5.47
    \end{tabular}
    \caption{Comparison of the execution time between different programming languages using the computer language benchmarks game\footnote{\url{https://benchmarksgame-team.pages.debian.net/benchmarksgame/index.html}}}
    \label{fig:clbg}
\end{table}

\subsection{Goal}

To reduce the energy consumption of Python, we started by targeting the main usage of this programming language, which is revealed to be data science and web development. Figure~\ref{fig:usecase} illustrates a study made by the JetBrain company on Python developers. \footnote{\fnurl{https://www.jetbrains.com/lp/python-developers-survey-2020}}.In this survey, multiple answers were accepted. As one can see, 57\% of the respondents said that they use Python for data science, 51\% said they are using it for web development, and around 40\% said they are using it for system administration.
\begin{figure}[hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_use_cases}
    \caption{Use cases of python }
    \label{fig:usecase}
\end{figure}

In this chapter, we will first investigate its behavior in the most common usage cases.Following that, we will focus on the language's structures in the hope of discovering some fundamental guidelines for more general purposes, as \citeauthor{hasan_energy_2016-1} did in their article~\cite{hasan_energy_2016-1}.
Following that, we will measure the energy consumption of several Python implementations in order to determine a non-intrusive technique to improve energy efficiency.

Therefore this chapter will target the following research questions :
\begin{compactenum}[\indent\bf RQ\,1:]
    \item \emph{What is the behavior of Python when it is used as a data science tool?}
    \item \emph{Are the default guidelines of Python energy efficient?}
    \item \emph{Can we reduce the energy consumption of Python programs without impacting the source code?}
\end{compactenum}
